---
title: "CrossFit Open 24.1 Statistics Recap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, warning=FALSE, message = FALSE, 
  include=TRUE, cache=TRUE)
```


```{r cars, eval=FALSE}
queryReq <- "https://c3po.crossfit.com/api/leaderboards/v2/competitions/open/2024/leaderboards?view=0&division=%d&region=0&scaled=%d&sort=0&page=%s"

numberOfPagesDf <- expand.grid(
  division=seq_len(2), ## 1=men; 2=women, one for each gender
  scaled=seq_len(3)-1 ## encoded 0=rx, 1=scaled, 2=foundations
  )


numberOfPagesDf$numPages <- vapply(seq_len(nrow(numberOfPagesDf)), function(i){
  jsonData <- 
    jsonlite::fromJSON(
      sprintf(queryReq, 
              numberOfPagesDf[i,"division"], numberOfPagesDf[i,"scaled"], 1))
  Sys.sleep(0.2)
  jsonData$pagination$totalPages
}, numeric(1))

allLeaderboard <- lapply(seq_len(nrow(numberOfPagesDf)), function(i){
  div <- numberOfPagesDf[i,"division"]
  sc <- numberOfPagesDf[i,"scaled"]
  pn <- numberOfPagesDf[i,"numPages"]
  pn <- seq_len(pn)
  set.seed(2)
  pn <- sample(pn, 28)
  tst <- lapply( pn, function(pageNum){
    tmpFile <- sprintf("cache/20240304openlead_div%d_sc%d_page%d.rds", div, sc, pageNum)
    if( !file.exists(tmpFile) ){
      cat(sprintf("div %d, sc %d, pg %d\n", div, sc, pageNum))
      dat <- 
        jsonlite::fromJSON(sprintf(queryReq, div, sc, pageNum))
      participants <- dat$leaderboardRows$entrant[,c("competitorId", "competitorName", "gender", "countryOfOriginCode", "affiliateName", "age", "height", "weight")]
      participantScores <- lapply(seq_len(nrow(participants)), function(x){
        compId <- dat$leaderboardRows$entrant$competitorId[x]
        scores <- dat$leaderboardRows$scores[[x]]$scoreDisplay
        retDat <- cbind( compId, Reduce(cbind, scores) )
        colnames(retDat) <- c("competitorId", "24.1")
        retDat
      })
      participantScores <- Reduce(rbind, participantScores)
      participantScores <- merge(participants, participantScores)
      saveRDS(participantScores, file=tmpFile)
      Sys.sleep(runif(1, min=1, max=5))
      participantScores
    }else{
      participantScores <- readRDS(tmpFile)
    }
    participantScores
  })
  tst <- dplyr::bind_rows(tst)
  tst$division <- div
  tst$scaled <- sc
  tst
})

allLeaderboard <- dplyr::bind_rows(allLeaderboard)

```


```{r, eval=FALSE}
library(lubridate)

allLeaderboardClean <- allLeaderboard |>
  dplyr::mutate( 
    `24.1`=ifelse(!grepl("\\d+:\\d+|reps", `24.1`), NA, `24.1`),
    division=ifelse( division == 1, "M", "F" ),
    ) |>
  tidyr::pivot_longer(
    cols=dplyr::matches("24."), 
    names_to="workout",
    values_to="score" ) |>
  dplyr::mutate(
    workoutType=dplyr::case_when(
      grepl("- s", score) ~ "scaled",
      grepl("- f", score) ~ "foundations",
      is.na(score) ~ NA_character_,
      TRUE ~ "Rx" ),
    score=gsub(" - s| - f", "" , score ),
    time=dplyr::case_when(
      workout == "24.1" & grepl("reps", score) ~ "15:00",
      workout == "24.1" & !grepl("reps", score) ~  score, 
      TRUE ~ NA),
    timeInSecs=ms(time),
    reps=dplyr::case_when(
      is.na(score) ~ NA,
      workout == "24.1" & grepl("reps", score) ~ gsub("(\\d+) reps", "\\1", score),
      workout == "24.1" & !grepl("reps", score) ~ "180",
      TRUE ~ NA),
    reps=as.numeric(reps),
    weightNumeric=dplyr::case_when(
      grepl("kg", weight) ~ as.numeric(gsub("(\\d+) kg", "\\1", weight)),
      grepl("lb", weight) ~ as.numeric(gsub("(\\d+) lb", "\\1", weight))*0.453592,
      weight == "" ~ NA_integer_
    ),
    weightNumeric=dplyr::case_when(
      grepl("kg", weight) ~ as.numeric(gsub("(\\d+) kg", "\\1", weight)),
      grepl("lb", weight) ~ as.numeric(gsub("(\\d+) lb", "\\1", weight))*0.453592,
      weight == "" ~ NA_integer_
    ),
    heightNumeric=dplyr::case_when(
      grepl("cm", height) ~ as.numeric(gsub("(\\d+) cm", "\\1", height)),
      grepl("in", height) ~ as.numeric(gsub("(\\d+) in", "\\1", height))*2.54,
      height == "" ~ NA_integer_
    ),
    age=as.numeric(age),
    ageGroup=cut( age, c(14, 16, 18, 35, 40, 45, 50, 55, 60, 65, Inf)-0.1)
  ) |>
  unique() |>
  dplyr::mutate( 
    affiliateNameShort=gsub("CrossFit | CrossFit", "", affiliateName ))

levels(allLeaderboardClean$ageGroup) <- 
  c("14-15", "16-17", "18-34", "35-39", "40-44", "45-49", "50-54", "55-59", 
    "60-64", ">65")

allLeaderboardClean |>
  split(allLeaderboardClean$ageGroup) |>
  purrr::map(\(df) range(df$age))

any(is.na(allLeaderboardClean$weightNumeric) & 
  allLeaderboardClean$weight != "")

any(is.na(allLeaderboardClean$heightNumeric) & 
  allLeaderboardClean$height != "")

saveRDS(allLeaderboardClean, file="data/allleaderboardClean241.rds")

```

The CrossFit Open is the largest participatory sporting event on Earth, 
according to the official website. In the first out of the three workouts, 
participants needed to perform the following series of exercise in a maximum
of 15 minutes,

- 21 Dumbbell Snatches (arm 1) (50/35 lb)
- 21 Lateral Burpees Over Dumbbell
- 21 Dumbbell Snatches (arm 2) (50/35 lb)
- 21 Lateral Burpees Over Dumbbell
- 15 Dumbbell Snatches (arm 1) (50/35 lb)
- 15 Lateral Burpees Over Dumbbell
- 15 Dumbbell Snatches(arm 2) (50/35 lb)
- 15 Lateral Burpees Over Dumbbell
- 9 Dumbbell Snatches (arm 1) (50/35 lb)
- 9 Lateral Burpees Over Dumbbell
- 9 Dumbbell Snatches (arm 2) (50/35 lb)
- 9 Lateral Burpees Over Dumbbell

`r ` participants submitted a score, of which `` participated in the Rx 
category, `` participated in the scaled category and `` participated in 
the foundations category. 



```{r}

library(ggplot2)
library(cowplot)
## library(plotly)

allLeaderboardClean <- readRDS("data/allleaderboardClean241.rds")
allLeaderboardClean <- allLeaderboardClean[which(!is.na(allLeaderboardClean$score)),]

allLeaderboardClean$workoutType <- 
  factor(allLeaderboardClean$workoutType, levels=c("Rx", "scaled", "foundations"))

pl <- allLeaderboardClean |>
  dplyr::filter(workout == "24.1") |>
  ggplot(aes(as.numeric(timeInSecs)/60)) +
  geom_histogram( ) +
  facet_grid(division~workoutType) +
  theme_cowplot() +
  panel_border(colour="black") +
  theme(axis.line=element_blank()) +
  labs(y="#of participants", x="Time")

pl
```


```{r}

allLeaderboardClean |>
  dplyr::filter( workout == "24.1", workoutType == "Rx" ) |>
  dplyr::group_by( division, ageGroup ) |>
  dplyr::summarize( cut25=quantile(as.numeric(timeInSecs), 0.25)/60 )

```


```{r, fig.height=5.6, fig.width=5, fig.res=300}


extraInclude <- data.frame(
  c("HelvetiX", "Basel", "Minic"),
  "Basel Box")
colnames(extraInclude) <- c("affiliateNameShort", "cat")

boxplotTopGyms <- function(  allLeaderboardClean, wod="24.1", rankVar="affiliateName", 
                             div="M", wodType="Rx", extraInclude=NULL, minParticipants=1, 
                             topN=10 ){
  topBoxes <- allLeaderboardClean |>
    dplyr::filter( workout == wod, workoutType == wodType, division == div ) |>
    dplyr::group_by( get(rankVar) ) |>
    dplyr::summarise( 
      meanTime=median(as.numeric(timeInSecs), na.rm=TRUE), 
      numparticipants=dplyr::n() ) |>
    dplyr::arrange( meanTime ) |>
    dplyr::filter( numparticipants > minParticipants ) |>
    head(topN) |>
    dplyr::select( `get(rankVar)` ) |>
    dplyr::mutate( cat="Top 10" )
  colnames(topBoxes) <- c(rankVar, "cat")
  topBoxes <- rbind(topBoxes, extraInclude)
  topLeaderboard <- allLeaderboardClean |>
    dplyr::filter( division == div, workoutType == wodType, division == div ) |>
    dplyr::select( competitorName, !!rankVar, timeInSecs ) |>
    dplyr::left_join( topBoxes ) |>
    dplyr::mutate( cat=ifelse( is.na(cat), "Other", cat ) )
   topLeaderboard[[rankVar]] <- 
     ifelse( topLeaderboard[[rankVar]] %in% topBoxes[[rankVar]], 
             topLeaderboard[[rankVar]], "Other" )
   topLeaderboard[[rankVar]] <- forcats::fct_reorder(
     topLeaderboard[[rankVar]], 
     as.numeric(topLeaderboard$timeInSecs),
     median)
   pl <- topLeaderboard |>
     ggplot( aes( get(rankVar), as.numeric(timeInSecs)/60, col=cat)) +
     geom_boxplot() +
     theme_cowplot() +
     theme( axis.text.x=element_text(angle=35, hjust=1) ) +
     labs(x="", y="minutes", col="",
          title=sprintf("%s %s %s", wod, div, wodType ))
   pl
}

allPossibilities <- expand.grid( 
  div=c("M", "F"), 
  wodType=c("Rx", "scaled", "foundations") )

allBoxplots <- lapply(seq_len(nrow(allPossibilities)), function(i){
  boxplotTopGyms( 
    allLeaderboardClean, 
    div=allPossibilities[i,"div"], extraInclude=extraInclude, 
    wodType=allPossibilities[i,"wodType"], 
    rankVar="affiliateNameShort" ) +
    theme(legend.pos="none") +
    coord_cartesian( ylim=c(5, 15) )
})

plots <- plot_grid( plotlist=allBoxplots, ncol=2, align="hv" )

legend_b <- get_legend(
  allBoxplots[[6]] + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "top"))

plot_grid(legend_b, plots, ncol = 1, rel_heights = c(0.1, 1))

```
