---
title: "CrossFit Open 24.1 Statistics Recap"
author: "Alejandro Reyes"
output: 
  bookdown::html_document2: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE, warning=FALSE, message = FALSE, 
  include=TRUE, cache=TRUE)
```


```{r cars, eval=FALSE}
queryReq <- "https://c3po.crossfit.com/api/leaderboards/v2/competitions/open/2024/leaderboards?view=0&division=%d&region=0&scaled=%d&sort=0&page=%s"

numberOfPagesDf <- expand.grid(
  division=seq_len(2), ## 1=men; 2=women, one for each gender
  scaled=seq_len(3)-1 ## encoded 0=rx, 1=scaled, 2=foundations
  )


numberOfPagesDf$numPages <- vapply(seq_len(nrow(numberOfPagesDf)), function(i){
  jsonData <- 
    jsonlite::fromJSON(
      sprintf(queryReq, 
              numberOfPagesDf[i,"division"], numberOfPagesDf[i,"scaled"], 1))
  Sys.sleep(0.2)
  jsonData$pagination$totalPages
}, numeric(1))

allLeaderboard <- lapply(seq_len(nrow(numberOfPagesDf)), function(i){
  div <- numberOfPagesDf[i,"division"]
  sc <- numberOfPagesDf[i,"scaled"]
  pn <- numberOfPagesDf[i,"numPages"]
  pn <- seq_len(pn)
  tst <- lapply( pn, function(pageNum){
    tmpFile <- sprintf("cache/20240304openlead_div%d_sc%d_page%d.rds", div, sc, pageNum)
    if( !file.exists(tmpFile) ){
      cat(sprintf("div %d, sc %d, pg %d\n", div, sc, pageNum))
      dat <- 
        jsonlite::fromJSON(sprintf(queryReq, div, sc, pageNum))
      participants <- dat$leaderboardRows$entrant[,c("competitorId", "competitorName", "gender", "countryOfOriginCode", "affiliateName", "age", "height", "weight")]
      participantScores <- lapply(seq_len(nrow(participants)), function(x){
        compId <- dat$leaderboardRows$entrant$competitorId[x]
        scores <- dat$leaderboardRows$scores[[x]]$scoreDisplay
        retDat <- cbind( compId, Reduce(cbind, scores) )
        colnames(retDat) <- c("competitorId", "24.1")
        retDat
      })
      participantScores <- Reduce(rbind, participantScores)
      participantScores <- merge(participants, participantScores)
      saveRDS(participantScores, file=tmpFile)
      Sys.sleep(0.75)
      participantScores
    }else{
      participantScores <- readRDS(tmpFile)
    }
    participantScores
  })
  tst <- dplyr::bind_rows(tst)
  tst$division <- div
  tst$scaled <- sc
  tst
})

allLeaderboard <- dplyr::bind_rows(allLeaderboard)
saveRDS(allLeaderboard, file="cache/allLeaderboardDF.rds")

```


```{r, eval=FALSE}
library(lubridate)

allLeaderboard <- readRDS("cache/allLeaderboardDF.rds")

allLeaderboardClean <- allLeaderboard |>
  dplyr::mutate( 
    `24.1`=ifelse(!grepl("\\d+:\\d+|reps", `24.1`), NA, `24.1`),
    division=ifelse( division == 1, "M", "F" ),
    ) |>
  tidyr::pivot_longer(
    cols=dplyr::matches("24."), 
    names_to="workout",
    values_to="score" ) |>
  dplyr::filter( !is.na(score) ) |>
  dplyr::mutate(
    workoutType=factor(dplyr::case_when(
      grepl("- s", score) ~ "scaled",
      grepl("- f", score) ~ "foundations",
      is.na(score) ~ NA_character_,
      TRUE ~ "Rx" ),levels=c("Rx", "scaled", "foundations")),
    score=gsub(" - s| - f", "" , score ),
    time=dplyr::case_when(
      workout == "24.1" & grepl("reps", score) ~ "15:00",
      workout == "24.1" & !grepl("reps", score) ~  score, 
      TRUE ~ NA),
    timeInSecs=ms(time),
    reps=dplyr::case_when(
      is.na(score) ~ NA,
      workout == "24.1" & grepl("reps", score) ~ gsub("(\\d+) reps", "\\1", score),
      workout == "24.1" & !grepl("reps", score) ~ "180",
      TRUE ~ NA),
    reps=as.numeric(reps),
    weightNumeric=dplyr::case_when(
      grepl("kg", weight) ~ as.numeric(gsub("(\\d+) kg", "\\1", weight)),
      grepl("lb", weight) ~ as.numeric(gsub("(\\d+) lb", "\\1", weight))*0.453592,
      weight == "" ~ NA_integer_
    ),
    weightNumeric=dplyr::case_when(
      grepl("kg", weight) ~ as.numeric(gsub("(\\d+) kg", "\\1", weight)),
      grepl("lb", weight) ~ as.numeric(gsub("(\\d+) lb", "\\1", weight))*0.453592,
      weight == "" ~ NA_integer_
    ),
    heightNumeric=dplyr::case_when(
      grepl("cm", height) ~ as.numeric(gsub("(\\d+) cm", "\\1", height)),
      grepl("in", height) ~ as.numeric(gsub("(\\d+) in", "\\1", height))*2.54,
      height == "" ~ NA_integer_
    ),
    age=as.numeric(age),
    ageGroup=cut( age, c(14, 16, 18, 35, 40, 45, 50, 55, 60, 65, Inf)-0.1)
  ) |>
  unique() |>
  dplyr::mutate( 
    affiliateNameShort=gsub("CrossFit", "CF", affiliateName ))

levels(allLeaderboardClean$ageGroup) <- 
  c("14-15", "16-17", "18-34", "35-39", "40-44", "45-49", "50-54", "55-59", 
    "60-64", ">65")

allLeaderboardClean |>
  split(allLeaderboardClean$ageGroup) |>
  purrr::map(\(df) range(df$age))

any(is.na(allLeaderboardClean$weightNumeric) & 
  allLeaderboardClean$weight != "")

any(is.na(allLeaderboardClean$heightNumeric) & 
  allLeaderboardClean$height != "")

saveRDS(allLeaderboardClean, file="data/allleaderboardClean241.rds")

```




```{r}

library(ggplot2)
library(cowplot)
## library(plotly)

allLeaderboardClean <- readRDS("data/allleaderboardClean241.rds")

##allLeaderboardClean <- allLeaderboardClean[which(!is.na(allLeaderboardClean$score)),]
##allLeaderboardClean$workoutType <- 
#3  factor(allLeaderboardClean$workoutType, levels=c("Rx", "scaled", "foundations"))
```


The CrossFit Open is the largest participatory sport event on Earth, 
according to the official website, and it consists of three workouts 
spread over three weeks. The first workout of the CrossFit Open 2024 
took place last weekend and participants needed to perform the following 
exercises within maximum 15 minutes:

- 21 Dumbbell Snatches (arm 1) (50/35 lb)
- 21 Lateral Burpees Over Dumbbell
- 21 Dumbbell Snatches (arm 2) (50/35 lb)
- 21 Lateral Burpees Over Dumbbell
- 15 Dumbbell Snatches (arm 1) (50/35 lb)
- 15 Lateral Burpees Over Dumbbell
- 15 Dumbbell Snatches(arm 2) (50/35 lb)
- 15 Lateral Burpees Over Dumbbell
- 9 Dumbbell Snatches (arm 1) (50/35 lb)
- 9 Lateral Burpees Over Dumbbell
- 9 Dumbbell Snatches (arm 2) (50/35 lb)
- 9 Lateral Burpees Over Dumbbell


## Global distribution of 24.1 scores {-}

An impressive `r nrow(allLeaderboardClean)` participants submitted a score 
to the CrossFit Open website: `r round(100*sum(allLeaderboardClean$workoutType == "Rx")/nrow(allLeaderboardClean))`% of participants competed in the "Rx" category, `r round(100*sum(allLeaderboardClean$workoutType == "scaled")/nrow(allLeaderboardClean))`% 
competed in the "scaled" category and `r round(100*sum(allLeaderboardClean$workoutType == "foundations")/nrow(allLeaderboardClean))`% competed in the "foundations" 
category. 


```{r global-dist, fig.width=5, fig.height=5, fig.cap="Distribution of completion times for 24.1."}

pl <- allLeaderboardClean |>
  dplyr::filter(workout == "24.1") |>
  dplyr::filter(workoutType %in% c("Rx", "scaled")) |>
  ggplot(aes(as.numeric(timeInSecs)/60)) +
  geom_histogram( ) +
  facet_grid(division~workoutType) +
  theme_cowplot() +
  panel_border(colour="black") +
  theme(axis.line=element_blank()) +
  labs(
    y="#of participants", x="Completion time (minutes)",
    caption="Alejandro Reyes. https://github.com/areyesq89/CFOpenStats.", 
    title="Completion times of CF Open 24.1")

pr <- round(100*sum(allLeaderboardClean$timeInSecs < (10*60))/nrow(allLeaderboardClean))

pl
```

Figure \@ref(fig:global-dist) shows the distribution of completion times
for each category for male and female divisions. Across all 
categories, only  
`r pr`%
of athletes completed the workout in less than 10 minutes and `r round(100*sum(allLeaderboardClean$reps < 180)/nrow(allLeaderboardClean))`% could not complete all the repetitions
within the 15 minute mark. It was a very though workout!

## How many seconds away are you from quarter-final scores? {-}

Athletes that rank among the top 25% in the Rx category qualify to the quarterfinals. 
Table \@ref(fig:global-dist) shows 
how fast an athlete would have had to complete 24.1 in order to be 
within the top 25% in their age category. 

```{r thetablemale, results='asis'}

allLeaderboardClean |>
  dplyr::filter( workout == "24.1", workoutType == "Rx", division =="M" ) |>
  dplyr::group_by( division, ageGroup ) |>
  dplyr::summarize( 
    `24.1 time`=seconds_to_period(quantile(as.numeric(timeInSecs), 0.25))) |>
  dplyr::rename( `Age group`=ageGroup ) |>
  dplyr::select( `Age group`, `24.1 time` ) |>
  knitr::kable(caption = "Time cap of the 25% fastest in 24.1 (male)") |>
  kableExtra::kable_styling(full_width = F)

```

```{r thetablefemale, results='asis'}

allLeaderboardClean |>
  dplyr::filter( workout == "24.1", workoutType == "Rx", division == "F") |>
  dplyr::group_by( division, ageGroup ) |>
  dplyr::summarize( 
    `24.1 time`=seconds_to_period(quantile(as.numeric(timeInSecs), 0.25))) |>
  dplyr::rename( `Age group`=ageGroup ) |>
  dplyr::select( `Age group`, `24.1 time` ) |>
  knitr::kable(caption = "Time cap of the 25% fastest in 24.1 (female)") |>
  kableExtra::kable_styling(full_width = F)

```

For example, I am a 35-year-old male and I completed
the Rx workout in 13:43, meaning that I'm more than two minutes away from being
a quarter-final athlete!

## Comparing scores between CF boxes {-}

Figure \@ref(fig:box) shows a comparison of completion times between: 
- CF boxes with at least ten athletes in the CF Open leaderboard that had the best average scores worldwide (blue) 
- CF boxes in Basel, where I live and work out (red)
- CF boxes in the rest of the world (green)

```{r  box, fig.height=6, fig.width=6, fig.res=300, fig.cap="Distribution of completion times of top 8 world boxes and basel boxes."}

extraInclude <- data.frame(
  c("CF HelvetiX", "CF Basel", "CF Basel Downtown", "CF Minic"),
  "Basel Box")
colnames(extraInclude) <- c("affiliateNameShort", "cat")

boxplotTopGyms <- function(  allLeaderboardClean, wod="24.1", rankVar="affiliateName", 
                             div="M", wodType="Rx", extraInclude=NULL, 
                             minParticipants=10, 
                             topN=8 ){
  topBoxes <- allLeaderboardClean |>
    dplyr::filter( workout == wod, workoutType == wodType, division == div ) |>
    dplyr::group_by( get(rankVar) ) |>
    dplyr::summarise( 
      meanTime=median(as.numeric(timeInSecs), na.rm=TRUE), 
      numparticipants=dplyr::n() ) |>
    dplyr::arrange( meanTime ) |>
    dplyr::filter( numparticipants > minParticipants ) |>
    head(topN) |>
    dplyr::select( `get(rankVar)` ) |>
    dplyr::mutate( cat=sprintf("World top %s",topN) )
  colnames(topBoxes) <- c(rankVar, "cat")
  topBoxes <- rbind(topBoxes, extraInclude)
  topLeaderboard <- allLeaderboardClean |>
    dplyr::filter( division == div, workoutType == wodType, division == div ) |>
    dplyr::select( competitorName, !!rankVar, timeInSecs ) |>
    dplyr::left_join( topBoxes ) |>
    dplyr::mutate( cat=ifelse( is.na(cat), "Rest of the world", cat ) )
   topLeaderboard[[rankVar]] <- 
     ifelse( topLeaderboard[[rankVar]] %in% topBoxes[[rankVar]], 
             topLeaderboard[[rankVar]], "Rest of the world" )
   topLeaderboard[[rankVar]] <- forcats::fct_reorder(
     topLeaderboard[[rankVar]], 
     as.numeric(topLeaderboard$timeInSecs),
     median)
   pl <- topLeaderboard |>
     ggplot( aes( get(rankVar), as.numeric(timeInSecs)/60, col=cat)) +
     geom_boxplot() +
     theme_cowplot() +
     theme( axis.text.x=element_text(angle=35, hjust=1) ) +
     labs(x="", y="minutes", col="",
          title=sprintf("%s %s %s", wod, div, wodType ))
   pl
}

allPossibilities <- expand.grid( 
  div=c("M", "F"), 
  wodType=c("Rx", "scaled", "foundations") )

allPossibilities <- expand.grid( 
  wodType=c("Rx"),
  div=c("M", "F") )

allBoxplots <- lapply(seq_len(nrow(allPossibilities)), function(i){
  boxplotTopGyms( 
    allLeaderboardClean, 
    div=as.character(allPossibilities[i,"div"]), extraInclude=extraInclude, 
    wodType=as.character(allPossibilities[i,"wodType"]), 
    rankVar="affiliateNameShort" ) +
    theme(legend.pos="none") +
    coord_cartesian( ylim=c(5, 15) )
})

plots <- plot_grid( plotlist=allBoxplots, ncol=1, align="hv" )

legend_b <- get_legend(
  allBoxplots[[1]] + 
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "top"))

plot_grid(legend_b, plots, ncol = 1, rel_heights = c(0.1, 1)) +
  draw_figure_label(label = "A. Reyes. https://github.com/areyesq89/CFOpenStats.",
                    position="bottom.right") +
    draw_figure_label(label = "Comparison of 24.1 between CF boxes",
                    position="top.left")

```

It is interesting that in the Rx category, the 
completion times of male athletes from my CF box (CF HelvetiX) are very similar to those of CF boxes across the rest of the world. However, the 
completion times of female athletes from CF HelvetiX are much closer to the world's top
eight boxes.

## Conclusion {-}

If you have to bet for someone in Basel to qualify to the CF Games quarter finals,
bet on a CF HelvetiX female athlete!